Floquet-Drude Conductivity in Dressed Quantum Hall Systems

### 2021-07-01 paper/v1.0/sections/2_schrodinger_problem.tex
Add section 2 of paper
Add part2 of Appendix B paper

Using Landau gauge for the stationary magnetic field, we can represent it using vector potential as $\vb{A}_{s} = (-By,0,0)^{\text{T}}$ and choosing Coulomb gauge, the vector potential of the dynamic dressing radiation can be presented as $\vb{A}_{d}(t) = (0,[E/\omega ]\cos(\omega t),0)^{\text{T}}$. These vector potentials are coupled to the momentum of 2DEG as kinetic momentum \cite{mahan81,bruus04} and this leads to the time-dependent Hamiltonian
\begin{equation} \label{eq_1}
  \hat{H}_e(t) = \frac{1}{2m_e}\Big[\hat{\vb{p}} - e\big(\vb{A}_{s}+\vb{A}_{d}(t)\big)\Big]^2,
\end{equation}
where $m_e$ is the effective electron mass and $e$ is the magnitude of the electron charge. $\hat{\vb{p}} = (\hat{p}_x,\hat{p}_y,0)^{\text{T}}$ represents the canonical momentum operator for 2DEG with electron momentums $p_{x,y}$.
The exact solutions for the time-dependent Schrödinger equation $i\hbar \dv{t}\psi = \hat{H}_e(t) \psi$ was already given by Refs. \cite{husmi53,ditt98,dini16} and we can present them as a set of wave functions defined by two quantum numbers $(n,m)$
\begin{equation} \label{eq_2}
  \begin{aligned}
    \psi_{n,m}&(x,y,t)  = \frac{1}{\sqrt{L_x}}
    \chi_n\left[y - y_0 - \zeta(t)\right]
    \text{exp}\bigg(
    \frac{i}{\hbar}\bigg[- \varepsilon_nt \\
    &
    + p_x x + \frac{eE(y - y_0)}{\omega}\cos(\omega t)+
    m_e\dot{\zeta}(t)\big[y - y_0 -\zeta(t)\big]\\
    & +
    \int_0^{t}dt'L(\zeta,\dot{\zeta},t')\bigg]\bigg),
  \end{aligned}
\end{equation}
where $n \in \mathbb{Z}^{+}_0$ and $m \in \mathbb{Z}$ ; see Appendix A. Here $L_{x,y}$ are dimention of the 2DEG surface, $\hbar$ is the reduced Planck constant, and $y_0 = -p_x/eB$ is the center of the cyclotron orbit along $y$ axis. $\chi_n$ are well known solutions(Gauss-Hermite functions) for Schrödinger equation of a stationary quantum harmonic oscillartor
\begin{equation} \label{eq_3}
  \chi_n(x) \equiv
   \frac{\sqrt{\kappa}}{\sqrt{2^{n}n!}}
  e^{-\kappa^2 x^2/2}
  \mathcal{H}_n \qty(\kappa x) \quad \text{with}
  \quad
  \kappa = \sqrt{\frac{m_e \omega_0}{\hbar}},
\end{equation}
with eigenvalues given by $\varepsilon_n = \hbar \omega_0 (n + 1/2)$ and $\omega_0 = eB/m_e$ is the cyclotron frequency. Each $n$ value defines the  energy($\varepsilon_n$) of the respective Landau level. The path shift of the driven classical oscillator $\zeta(t)$ is given by
\begin{equation} \label{eq_4}
  \zeta(t) = \frac{eE}{m_e(\omega_0^2 - \omega^2)}\sin(\omega t),
\end{equation}
while the Lagrangian of the classical oscialltor $L(\zeta,\dot{\zeta},t)$ can be idenfied as
\begin{equation} \label{eq_5}
  L(\zeta,\dot{\zeta},t) = \frac{1}{2} m_e\dot{\zeta}^2(t) - \frac{1}{2}m_e\omega_0^2 \zeta^2(t) + eE\zeta(t) \sin(\omega t).
\end{equation}

x

>>>>

Using Landau gauge for the stationary magnetic field, we can represent it using vector potential as $\vb{A}_{s} = (-By,0,0)^{\text{T}}$ and choosing Coulomb gauge, the vector potential of the dynamic dressing radiation can be presented as $\vb{A}_{d}(t) = (0,[E/\omega ]\cos(\omega t),0)^{\text{T}}$. These vector potentials are coupled to the momentum of 2DEG as kinetic momentum \cite{mahan81,bruus04} and this leads to the time-dependent Hamiltonian
\begin{equation} \label{eq_1}
  \hat{H}_e(t) = \frac{1}{2m_e}\Big[\hat{\vb{p}} - e\big(\vb{A}_{s}+\vb{A}_{d}(t)\big)\Big]^2,
\end{equation}
where $m_e$ is the effective electron mass and $e$ is the magnitude of the electron charge. $\hat{\vb{p}} = (\hat{p}_x,\hat{p}_y,0)^{\text{T}}$ represents the canonical momentum operator for 2DEG with electron momentums $p_{x,y}$.
The exact solutions for the time-dependent Schrödinger equation $i\hbar \dv{t}\psi = \hat{H}_e(t) \psi$ was already given by Refs. \cite{husmi53,ditt98,dini16} and we can present them as a set of wave functions defined by two quantum numbers $(n,m)$
\begin{equation} \label{eq_2}
  \begin{aligned}
    \psi_{n,m}&(x,y,t)  = \frac{1}{\sqrt{L_x}}
    \chi_n\left[y - y_0 - \zeta(t)\right]
    \text{exp}\bigg(
    \frac{i}{\hbar}\bigg[- \varepsilon_nt \\
    &
    + p_x x + \frac{eE(y - y_0)}{\omega}\cos(\omega t)+
    m_e\dot{\zeta}(t)\big[y - y_0 -\zeta(t)\big]\\
    & +
    \int_0^{t}dt'L(\zeta,\dot{\zeta},t')\bigg]\bigg),
  \end{aligned}
\end{equation}
where $n \in \mathbb{Z}^{+}_0$ and $m \in \mathbb{Z}$ ; see Appendix A. Here $L_{x,y}$ are dimention of the 2DEG surface, $\hbar$ is the reduced Planck constant, and $y_0 = -p_x/eB$ is the center of the cyclotron orbit along $y$ axis. $\chi_n$ are well known solutions for Schrödinger equation of a stationary quantum harmonic oscillartor
\begin{equation} \label{eq_3}
  \chi_n(x) \equiv
   \frac{\sqrt{\kappa}}{\sqrt{2^{n}n!}}
  e^{-\kappa^2 x^2/2}
  \mathcal{H}_n \qty(\kappa x) \quad \text{with}
  \quad
  \kappa = \sqrt{\frac{m_e \omega_0}{\hbar}},
\end{equation}
with eigenvalues given by $\varepsilon_n = \hbar \omega_0 (n + 1/2)$ and $\omega_0 = eB/m_e$ is the cyclotron frequency. Each $n$ value defines the  energy($\varepsilon_n$) of the respective Landau level. The path shift of the driven classical oscillator $\zeta(t)$ is given by
\begin{equation} \label{eq_4}
  \zeta(t) = \frac{eE}{m_e(\omega_0^2 - \omega^2)}\sin(\omega t),
\end{equation}
while the Lagrangian of the classical oscialltor $L(\zeta,\dot{\zeta},t)$ can be idenfied as
\begin{equation} \label{eq_5}
  L(\zeta,\dot{\zeta},t) = \frac{1}{2} m_e\dot{\zeta}^2(t) - \frac{1}{2}m_e\omega_0^2 \zeta^2(t) + eE\zeta(t) \sin(\omega t).
\end{equation}
The exponential phase shifts in Eq.~(\ref{eq_2}) represent the influence done by the stationary magnetic field and strong dressing field. Therefore we can accept that magnetotransport properties of 2DEG will be renormalized by the magnetic field as well as the dressing field.

### 2021-07-01 paper/v1.0/sections/appendix_a.tex
Add part2 of Appendix B paper

The deriving process of solutions for Schrödinger equation with Hamiltonian of an electron in 2DEG (Eq. \ref{eq_1}) quite similar to that followed in Refs. \cite{husmi53,dini16}. We start with expanding the Hamiltonian for two-dimentional case
\begin{equation} \label{eq_a1}
  \hat{H}_e(t) = \frac{1}{2m_e}\left[
    \left(\hat{p}_x + eBy \right)^2 +
    \left(\hat{p}_y - \frac{eE}{\omega}\cos(\omega t)\right)^2
  \right],
\end{equation}
and since $\left[\hat{H}_e(t),\hat{p}_x \right] =0$ both operators share same (simultaneous) eigen functions
$\frac{1}{\sqrt{L_x}}\exp(\frac{ip_x x}{\hbar})$ with $p_x = m2\pi \hbar/L_x~,~ m \in \mathbb{Z}$.
Therefore we re-arrange the Hamiltonian using definition of canonical momentum in $y$-direction to derive
\begin{equation} \label{eq_a2}
    \hat{H}_e(t) = \frac{1}{2m_e}\left[
      \left({p}_x + eBy \right)^2 +
      \left(-i\hbar \pdv{y}- \frac{eE}{\omega}\cos(\omega t)\right)^2
    \right].
\end{equation}
We now define the \textit{center of the cyclotron orbit} along $y$ axis $y_0 \equiv {-p_x}/{eB}$ and the \textit{cyclotron frequency} $\omega_0 \equiv {eB}/{m_e}$. This leads to a new arangement of the Hamiltonian
\begin{equation} \label{eq_a3}
  \begin{aligned}
    \hat{H}_e(t) =
      \frac{m_e \omega_0^2}{2}\tilde{y}^2 +
      \frac{1}{2m_e}\bigg[
      -\hbar^2 \pdv[2]{\tilde{y}} & +
      \frac{2i\hbar eE}{\omega}\cos(\omega t) \pdv{\tilde{y}} \\
      & +
      \frac{e^2E^2}{\omega^2}\cos[2](\omega t)
      \bigg],
  \end{aligned}
\end{equation}
where we used a variable substitution $\tilde{y} = (y - y_0)$. Now we are assuming that the solutions for the time-dependent Schrödinger equation
\begin{equation} \label{eq_a4}
    i \hbar \dv{\psi}{t} = \hat{H}_e(t)\psi,
\end{equation}
can present by the following form
\begin{equation} \label{eq_a5}
    \psi_m(x,\tilde{y},t) = \frac{1}{\sqrt{L_x}} \exp\bigg(
      \frac{ip_x x}{\hbar} +
      \frac{ieE\tilde{y}}{\hbar \omega}\cos(\omega t)
    \bigg) \vartheta(\tilde{y},t),
\end{equation}
where $\vartheta(\tilde{y},t)$ is a function that need to be find to satisfy the following property
\begin{equation} \label{eq_a6}
    \bigg[
    \frac{m_e \omega_0^2}{2}\tilde{y}^2
    - {eE\tilde{y}}\sin(\omega t)
    -
    \frac{\hbar^2}{2m_e}
    \pdv[2]{\tilde{y}}
    - i \hbar \dv{t}
    \bigg]
    \vartheta(\tilde{y},t) = 0.
\end{equation}
If we turn off the strong dressing field ($E=0$), this equation leads to simple harmonic oscillator Hamiltonian
\begin{equation} \label{eq_a7}
     i \hbar \dv{\vartheta(\tilde{y},t)}{t} =
    \bigg[
    \frac{\hat{p}_{\tilde{y}}^2}{2m_e} +
    \frac{1}{2}m_e \omega_0^2\tilde{y}^2
    \bigg]
    \vartheta(\tilde{y},t).
\end{equation}
It is important to notice that we can identify the $S(t) \equiv eE\sin(\omega t)$ part as a external force act on the harmonic oscillator and we can solve this as a forced harmonic oscillator in $\tilde{y}$ axis.
\begin{equation} \label{eq_a8}
  \begin{aligned}
    i \hbar \dv{\vartheta(\tilde{y},t)}{t} =
    \bigg[
    -
    \frac{\hbar^2}{2m_e}
    \pdv[2]{\tilde{y}} +
    \frac{1}{2}m_e \omega_0^2\tilde{y}^2
    - \tilde{y}S(t)]
    \bigg]
    \vartheta(\tilde{y},t).
  \end{aligned}
\end{equation}

>>>>

The deriving process of solutions for Schrödinger equation with Hamiltonian of an electron in 2DEG (Eq. \ref{eq_1}) quite similar to that followed in Refs. \cite{husmi53,dini16}. We start with expanding the Hamiltonian for two-dimentional case
\begin{equation} \label{eq_a1}
  \hat{H}_e(t) = \frac{1}{2m_e}\left[
    \left(\hat{p}_x + eBy \right)^2 +
    \left(\hat{p}_y - \frac{eE}{\omega}\cos(\omega t)\right)^2
  \right],
\end{equation}
and since $\left[\hat{H}_e(t),\hat{p}_x \right] =0$ both operators share same (simultaneous) eigen functions
$\frac{1}{\sqrt{L_x}}\exp(\frac{ip_x x}{\hbar})$ with $p_x = 2\pi \hbar m/L_x~,~ m \in \mathbb{Z}$.
Therefore we re-arrange the Hamiltonian using definition of canonical momentum in $y$-direction to derive
\begin{equation} \label{eq_a2}
    \hat{H}_e(t) = \frac{1}{2m_e}\left[
      \left({p}_x + eBy \right)^2 +
      \left(-i\hbar \pdv{y}- \frac{eE}{\omega}\cos(\omega t)\right)^2
    \right].
\end{equation}
We now define the \textit{center of the cyclotron orbit} along $y$ axis $y_0 \equiv {-p_x}/{eB}$ and the \textit{cyclotron frequency} $\omega_0 \equiv {eB}/{m_e}$. This leads to a new arangement of the Hamiltonian
\begin{equation} \label{eq_a3}
  \begin{aligned}
    \hat{H}_e(t) =
      \frac{m_e \omega_0^2}{2}\tilde{y}^2 +
      \frac{1}{2m_e}\bigg[
      -\hbar^2 \pdv[2]{\tilde{y}} & +
      \frac{2i\hbar eE}{\omega}\cos(\omega t) \pdv{\tilde{y}} \\
      & +
      \frac{e^2E^2}{\omega^2}\cos[2](\omega t)
      \bigg],
  \end{aligned}
\end{equation}
where we used a variable substitution $\tilde{y} = (y - y_0)$. Now we are assuming that the solutions for the time-dependent Schrödinger equation
\begin{equation} \label{eq_a4}
    i \hbar \dv{\psi}{t} = \hat{H}_e(t)\psi,
\end{equation}
can present by the following form
\begin{equation} \label{eq_a5}
    \psi_m(x,\tilde{y},t) = \frac{1}{\sqrt{L_x}} \exp\bigg(
      \frac{ip_x x}{\hbar} +
      \frac{ieE\tilde{y}}{\hbar \omega}\cos(\omega t)
    \bigg) \vartheta(\tilde{y},t),
\end{equation}
where $\vartheta(\tilde{y},t)$ is a function that need to be find to satisfy the following property
\begin{equation} \label{eq_a6}
    \bigg[
    \frac{m_e \omega_0^2}{2}\tilde{y}^2
    - {eE\tilde{y}}\sin(\omega t)
    -
    \frac{\hbar^2}{2m_e}
    \pdv[2]{\tilde{y}}
    - i \hbar \dv{t}
    \bigg]
    \vartheta(\tilde{y},t) = 0.
\end{equation}
If we turn off the strong dressing field ($E=0$), this equation leads to simple harmonic oscillator Hamiltonian
\begin{equation} \label{eq_a7}
     i \hbar \dv{\vartheta(\tilde{y},t)}{t} =
    \bigg[
    \frac{\hat{p}_{\tilde{y}}^2}{2m_e} +
    \frac{1}{2}m_e \omega_0^2\tilde{y}^2
    \bigg]
    \vartheta(\tilde{y},t).
\end{equation}
It is important to notice that we can identify the $S(t) \equiv eE\sin(\omega t)$ part as a external force act on the harmonic oscillator and we can solve this as a forced harmonic oscillator in $\tilde{y}$ axis.
\begin{equation} \label{eq_a8}
  \begin{aligned}
    i \hbar \dv{\vartheta(\tilde{y},t)}{t} =
    \bigg[
    -
    \frac{\hbar^2}{2m_e}
    \pdv[2]{\tilde{y}} +
    \frac{1}{2}m_e \omega_0^2\tilde{y}^2
    - \tilde{y}S(t)]
    \bigg]
    \vartheta(\tilde{y},t).
  \end{aligned}
\end{equation}

### 2021-07-01 paper/v1.0/sections/appendix_b.tex
Add part2 of Appendix B paper

@@

x

>>>>

\subsection{Position space representation}

@@

\subsection{Momentum space representation}

To write the Floquet modes in momentum space, we perform continuous Fourier transform over the considering confined space $A=L_xL_y$ for Eq.~(\ref{eq_7})
\begin{equation} \label{eq_b8}
  \begin{aligned}
    \phi_{n,m}(k_x,k_y,t)  &=
    \int_{-L_y/2}^{L_y/2} dy\; \exp(-i\qty[k_y - \gamma(t)]y)
    \chi_{n}\qty[y - \mu(t)] \\
     & \times
     \frac{1}{\sqrt{L_x}}
     \int_{-L_x/2}^{L_x/2} dx\;
     \exp(-ik_x x)
     \exp( \frac{i p_x }{\hbar}x ) \\
     &
     \times
     \exp(
      \frac{-i\gamma(t)}{\hbar}
      y_0)
     \exp(\frac{-i}{\hbar}
     \qty[
     m_e \dot{\zeta}(t) \zeta(t) - \xi
     ]),
  \end{aligned}
\end{equation}
where
\begin{equation} \label{eq_b9}
  \mu(t) \equiv \frac{eE\sin(\omega t)}{m_e(\omega_0^2 - \omega^2)} + y_0
  \quad,\quad
  \gamma(t) \equiv
  \frac{eE\omega_0^2\cos(\omega t)}{\hbar\omega(\omega_0^2 - \omega^2)}.
\end{equation}
Next using the identity \cite{bruus04}
\begin{equation} \label{eq_b10}
  \int_{L_x} dx\;
  \exp( -ik_x x + \frac{i p_x }{\hbar}x ) =
  L_x \delta_{k_x,\frac{p_x}{\hbar}},
\end{equation}
we can derive
\begin{equation} \label{eq_b11}
  \begin{aligned}
    \phi_{n,m}(k_x,k_y,t)  =
    \exp(
     \frac{-i\gamma(t)}{\hbar}
     y_0) &
    \exp(\frac{-i}{\hbar}
    \qty[
    m_e \dot{\zeta}(t) \zeta(t) - \xi
    ]) \\
    & \times
    \Phi_{n,m}(k_y,t)
    \delta_{k_x,\frac{p_x}{\hbar}}.
  \end{aligned}
\end{equation}
Here we defined $\Phi_{n,m}(k_y,t)$ as
\begin{equation} \label{eq_b12}
  \Phi_{n,m}(k_y,t) \equiv
  \sqrt{L_x}
  \int_{-L_y/2}^{L_y/2} dy\;
  \chi_{n}\qty[y - \mu(t)]
  \exp(
    -i\qty[k_y - \gamma(t)]
    y).
\end{equation}
Subtituting $  {k'_y} = k_y -\gamma(t)$ and $y' = y -\mu(t)$ and assuming that size of the 2DEG sample in $y$-direction is large ($L_y \rightarrow \infty$), we can obtain
\begin{equation} \label{eq_b13}
  \Phi_{n,m}({k'_y} ,t) =
  {\sqrt{L_x}} e^{-i {k'_y}\mu}
  \int_{-\infty}^{\infty} dy'\;
  \chi_{n}\qty(y')
  \exp(-i{k'_y} y').
\end{equation}
We can identify that the integral represnts the Fourier transform of $\{\chi_n\}$ functions and using the symmetric conditions \cite{bruus04} for the Fourier transform of Gauss-Hermite functions $\vartheta_n(x)$:
\begin{equation} \label{eq_b14}
  \mathcal{FT}[\vartheta_n(\kappa x),x,k] = \frac{i^n}{|\kappa|}\vartheta_n(k/\kappa),
\end{equation}
Eq.~(\ref{eq_b13}) can be simplified as
\begin{equation} \label{eq_b15}
  \Phi_{n,m}({k'_y} ,t) =
    \sqrt{L_x}e^{-i {k'_y}\mu}
    \tilde{\chi}_{n}\qty({k'_y}),
\end{equation}
with
\begin{equation} \label{eq_b16}
  \tilde{\chi}_{n}\qty(k) =
  \frac{i^n}{\sqrt{2^{n} n! \sqrt{\pi}}}
  \qty(\frac{1}{\kappa})^{1/2}
  e^{-\frac{k^2}{2 \kappa^2}}
  \mathcal{H}_{\alpha} \qty(\frac{k}{\kappa}).
\end{equation}
Subtitute Eq.~(\ref{eq_b15}) back in Eq.~(\ref{eq_b11}) and this leads to
\begin{equation} \label{eq_b17}
  \begin{aligned}
    \phi_{n,m}(k_x,k_y,t)  = &
    {\sqrt{L_x}}
    \tilde{\chi}_{n}\qty(k_y - b\cos(\omega t)) \\
    & \times
    \text{exp}\left(
      i\xi
      -ik_y  \qty[d\sin(\omega t) + \frac{\hbar k_x}{eB}]
    \right),
  \end{aligned}
\end{equation}
where
\begin{equation} \label{eq_b18}
  b \equiv
  \frac{eE\omega_0^2}{\hbar\omega(\omega_0^2 - \omega^2)} \quad
  d \equiv
 \frac{eE}{m_e(\omega_0^2 - \omega^2)},
\end{equation}
and it is necessary to notice that $k_x$ is quantized with $k_x = 2\pi m/L_x ~,~ m \in \mathbb{Z}$.
